---
title: "Hands-on Programming with R"
author: "Feng Ji"
date: "July 29, 2016"
output: html_document
---
```{r Foreword}
The author is very nice and noticed why programming with R is an essential part of being a qualified data scientist. 
>Using the functions in R is like riding
>a bus. Writing programs in R is like driving a car.
```
      


```{r Project 1 weighted dice Chap 1}
# some common functions
ls()
round()
#outer product and inner product
a %*% b
a %o% b

#sampling 
sample (x = 1:4, size = 2)


args()
example()
#default is without replacement
sample(x = 1: 4, size = 2, replace = TRUE)
#Casino Officer
Casino <- function(){
  sample(x = 1:6, size = 2, replace = T)
}
Casino()

>Sampling with replacement is an easy way to create independent random samples

#Constructor
Casino <- function(a, b = 6){
  sample(x = 1:b, size = a, replace = T)
}
Casino(6)

```

```{r Project 1 weighted dice Chap 2}
x <- c(-1, -0.8, -0.6, -0.4, -0.2, 0, 0.2, 0.4, 0.6, 0.8, 1)
y <- x^3
#Polymorphism
library(ggplot2)
qplot(x, y)

x <- c(1, 2, 2, 2, 3, 3)
qplot(x, binwidth = 1)

#replicate
rep(3, 20)
rep(Casino(5),5)


#replicate
replicate(5, Casino(5))

Casino(5)
rolls <- replicate(100000, Casino(5))
qplot(rolls)

#rolls
roll <- function() {
die <- 1:6
dice <- sample(die, size = 2, replace = TRUE)
sum(dice)
}

rolls <- replicate(1000, roll())
qplot(rolls, binwidth = 1)



roll1 <- function(){
  die <- 1:6
  dice <- sample(die, size = 2, replace = T, prob = biased)
  sum(dice)
}
example(sample)
?sample


#biased dice
biased <- c(1/8, 1/8, 1/8, 1/8, 1/8, 3/8)

roll1()
x <- replicate(10000, roll1())
qplot(x, binwidth = 1)



#project 1 done
```

```{r Project 2 Playing Cards Chap 3}
die <- c(1:6)
is.vector(die)
#six basic data types of atomatic vectors:
#double integers characters logicals complex raw
names(die)
names(die) <- c("one", "two", "three", "four", "five", "six")
attributes(die)
#get rid of it
names(die) <- NULL

die
#using `dim` can put it to a n dimentional matrix which is exciting
dim(die) <- c(2,3)
die

dim(die) <- c(3, 2)
die

dim(die) <- c(1, 2, 3)
die #Can be only show in this way given the two dimentional screen

m <- matrix(die, nrow = 2)
m#by default it go by cols

m <- matrix(die, nrow = 2, byrow = T)
m

ar <- array(c(11:14, 21:24, 31:34), dim = c(2, 2, 3))
ar

x <- c(11:14, 21:24, 31:34)
cardName <- c("ace", "king", "queen", "jack", "ten")
clusters <- c("spades", "spades", "spades", "spades", "spades")
x <- matrix(c(cardName, clusters), nrow = 5, ncol = 2)

class(die)
class(x)
attributes(x)

#dim attributes was added and will be used by other functions

now <- Sys.time()
now
#time is constructed by  POSIXct framework
class(now)
unclass(now)

#test time framework by adding POSIXct framework as an attribute
testTime <- 0 #this is the secs that you want to add into the  the Universal Time Coordinated (UTC) zone
testTime
class(testTime) <- c("POSIXct", "POSIXt")
testTime

#factor

gender <- factor(c("male", "female", "male", "female"))
typeof(gender)

attributes(gender)
unclass(gender)


str(gender)
class(gender)

gender

as.character(gender)

#making card
card <- c("ace", "heart", 1)
typeof(card)
class(card)

#Coercion
t <- c(T, T, F, T)
class(t)
unclass(t)
sum(t)

as.character(t)
as.numeric(t)
as.logical(t)

#list
l1 <- list(1:100, T, F, "chahaha", list(T, F))
l1
card <- list("ace", "hearts", 1)
card

#data frame
#data frame is a special list
df <- data.frame(face = c("ace", "two", "six"),
suit = c("clubs", "clubs", "clubs"), value = c(1, 2, 3))
xx <- df <- data.frame(face = c("ace", "two", "six"),
suit = c("clubs", "clubs", "clubs"), value = c(1, 2, 3))

class(xx)
str(df)

df <- data.frame(face = c("ace", "two", "six"), suit = c("clubs", "clubs", "clubs"), value = c(1, 2, 3), stringsAsFactors = F)

df
str(df)
deck <- data.frame(
face = c("king", "queen", "jack", "ten", "nine", "eight", "seven", "six",
"five", "four", "three", "two", "ace", "king", "queen", "jack", "ten",
"nine", "eight", "seven", "six", "five", "four", "three", "two", "ace",
"king", "queen", "jack", "ten", "nine", "eight", "seven", "six", "five",
"four", "three", "two", "ace", "king", "queen", "jack", "ten", "nine",
"eight", "seven", "six", "five", "four", "three", "two", "ace"),
suit = c("spades", "spades", "spades", "spades", "spades", "spades",
"spades", "spades", "spades", "spades", "spades", "spades", "spades",
"clubs", "clubs", "clubs", "clubs", "clubs", "clubs", "clubs", "clubs",
"clubs", "clubs", "clubs", "clubs", "clubs", "diamonds", "diamonds",
"diamonds", "diamonds", "diamonds", "diamonds", "diamonds", "diamonds",
"diamonds", "diamonds", "diamonds", "diamonds", "diamonds", "hearts",
"hearts", "hearts", "hearts", "hearts", "hearts", "hearts", "hearts",
"hearts", "hearts", "hearts", "hearts", "hearts"),
value = c(13, 12, 11, 10, 9, 8, 7, 6, 5, 4, 3, 2, 1, 13, 12, 11, 10, 9, 8,
7, 6, 5, 4, 3, 2, 1, 13, 12, 11, 10, 9, 8, 7, 6, 5, 4, 3, 2, 1, 13, 12, 11,
10, 9, 8, 7, 6, 5, 4, 3, 2, 1)
)

#head and tail also can be added arguments...I did not know this!
head(deck, 100)
tail(deck, 1)

getwd()
write.csv(deck, file = "cards.csv", row.names = FALSE)
> This will prevent R from
> adding a column of numbers at the start of your data frame. 

```

```{r Project 2 Playing Cards Chap 4}
deal <- function(x){
  n <- sample(1:52, 1)
  return(x[n,])
}
deal(deck)

shuffle <- function(x){
  random <- sample(1:52, 52)
  return(x[random,])
}

shuffle(deck)
```

```{r Project 2 Playing Cards Chap 5}
#notation
lst <- list(numbers = c(1, 2), logical = TRUE, strings = c("a", "b", "c"))
lst
lst[1]

deck2 <- deck
deck2$new <- 1:52
deck2

1 %in% c(3, 4, 5)
## FALSE
c(1, 2) %in% c(3, 4, 5)
## FALSE FALSE
c(1, 2, 3) %in% c(3, 4, 5)
## FALSE FALSE TRUE
c(1, 2, 3, 4) %in% c(3, 4, 5)
## FALSE FALSE TRUE TRUE



#* "%in% looks beautiful"*
deck$face
sum(deck$face == "ace")
deck3 <- deck
deck3$value[deck3$face == "ace"]
deck$value[deck3$face == "ace"] <- 14
head(deck,111)

#dplyr is way better than this original method
deck4 <- deck3
deck4$value[deck4$suit == "hearts"] <- 1


```


```{r Project 2 Playing Cards Chap 6}
#Environment
library(devtools)
as.environment("package:stats")
globalenv()
baseenv()
emptyenv()
parent.env(globalenv())
ls(emptyenv())
ls(globalenv())
ls(baseenv())

head(globalenv()$deck, 3)

assign("new", "Hello Global", envir = globalenv())
new

#list your active environment
environment()


show_env <- function(){
list(ran.in = environment(),
parent = parent.env(environment()),
objects = ls.str(environment()))
}

show_env()
ls.str()

show_env <- function(){
a <- 1
b <- 2
  list(ran.in = environment(),
parent = parent.env(environment()),
objects = ls.str(environment()))
}

environment(show_env)
show_env()

foo <- "take me to your runtime"
show_env <- function(x = foo){
list(ran.in = environment(),
parent = parent.env(environment()),
objects = ls.str(environment()))
}

m <- 3
addd <- function(x){
  a <- mm
  return(a + x)
}
addd(3)

deal <- function() {

card <- deck[1, ]
assign("deck", deck[-1, ], envir = globalenv())  
card
}

deal()

head(deck, 3)
a <- shuffle(deck)
head(a, 3)


shuffle <- function(){
random <- sample(1:52, size = 52)
assign("deck", DECK[random, ], envir = globalenv())
}

shuffle()

setup <- function(deck) {
DECK <- deck
DEAL <- function() {
card <- deck[1, ]
assign("deck", deck[-1, ], envir = globalenv())
card
}
SHUFFLE <- function(){
random <- sample(1:52, size = 52)
assign("deck", DECK[random, ], envir = globalenv())
}
list(deal = DEAL, shuffle = SHUFFLE)
}
DEAL


setup(deck)
show_env()



#**Important** *Closures*

setup <- function(deck) {
DECK <- deck
DEAL <- function() {
card <- deck[1, ]
assign("deck", deck[-1, ], envir = globalenv())
card
}
SHUFFLE <- function(){
random <- sample(1:52, size = 52)
assign("deck", DECK[random, ], envir = globalenv())
}
}


setup <- function(deck) {
DECK <- deck
DEAL <- function() {
card <- deck[1, ]
assign("deck", deck[-1, ], envir = parent.env(environment()))
card
}
SHUFFLE <- function(){
random <- sample(1:52, size = 52)
assign("deck", DECK[random, ], envir = parent.env(environment()))
}
list(deal = DEAL, shuffle = SHUFFLE)
}




setup
```


```{r Project 3 Slot Machine Chap7}
#the final aim is to write a function that can act like a slot machine
get_symbols <- function() {
  wheel <- c("DD", "7", "BBB", "BB", "B", "C", "0")
  sample(wheel, size = 3, replace = TRUE,
  prob = c(0.03, 0.03, 0.06, 0.1, 0.25, 0.01, 0.52))
}

symbols <- get_symbols()



#Strategy
1. Break complex tasks into simple subtasks.
2. Use concrete examples.
3. Describe your solutions in English, then convert them to R.

#Checking three of a kind
all(symbols == symbol[1])

length(unique(symbols == 1)
#unique function returns the unique element in a vector

if(all(symples == symbols[1])){
  prize <- #look up the prize
} else if (#case 2 was selected){
  else {
  #count cherries
  }
}

#Deal with a big prize
symbols <- c("B", "BBB", "BB")
#Test all symbols are bars
unique(symbols)
unique(c("AB", "BA", "ABB", "AB"))

#%in% is very easy to use
all(symbols%in%c("B", "BB", "BBB"))
symbols%in%c("B", "BB", "BBB")

bars <- all(symbols%in%c("B", "BB", "BBB"))

#lookup table
payouts <- c("DD" = 100, "7" = 80, "BBB" = 40, "BB" = 25,
"B" = 10, "C" = 10, "0" = 0)
payouts
payouts["DD"]
unname(payouts["DD"])
unname(payouts["DD"])

symbols[1]
payouts[symbols[1]]

any(symbols%in%c("C"))
sum(symbols == "C")
#the beauty of vectors 
#you can conduct a "loop" within one line of code

sum(symbols == "DD")#count the total numbers of Dimonds

#with R we can change all the "Swich" phraze in Java as subsetting, which is amazing
#for example
c(0, 2, 5)[1]


same <- symbols[1] == symbols[2] && symbols[2] == symbols[3]
bars <- symbols %in% c("B", "BB", "BBB")
if (same) {
payouts <- c("DD" = 100, "7" = 80, "BBB" = 40, "BB" = 25,
"B" = 10, "C" = 10, "0" = 0)
prize <- unname(payouts[symbols[1]])
} else if (all(bars)) {
prize <- 5
} else {
cherries <- sum(symbols == "C")
prize <- c(0, 2, 5)[cherries + 1]
}
diamonds <- sum(symbols == "DD")
# double the prize if necessary

>As a general
>rule, use an if tree if each branch of the tree runs different code. Use a lookup table if
>each branch of the tree only assigns a different value.


score <- function (symbols) {
# identify case
same <- symbols[1] == symbols[2] && symbols[2] == symbols[3]
bars <- symbols %in% c("B", "BB", "BBB")
# get prize
if (same) {
payouts <- c("DD" = 100, "7" = 80, "BBB" = 40, "BB" = 25,
"B" = 10, "C" = 10, "0" = 0)
prize <- unname(payouts[symbols[1]])
} else if (all(bars)) {
prize <- 5
} else {
cherries <- sum(symbols == "C")
prize <- c(0, 2, 5)[cherries + 1]
}
# adjust for diamonds
diamonds <- sum(symbols == "DD")
prize * 2 ^ diamonds
}

play <- function() {
symbols <- get_symbols()
print(symbols)
score(symbols)
}

play()

```

```{r Project 3 S3 Chap8}
#class
n <- 10000
class(n)

attributes(n)
attributes(play)
attributes(deck)

one_play <- play()
one_play
attributes(one_play)
attr(one_play, "symbols") <- c("B", "0", "B")
attr(one_play, "symbols")
attributes(one_play)
one_play+1

play <- function() {
symbols <- get_symbols()
prize <- score(symbols)
attr(prize, "symbols") <- symbols
prize
}
play()
n <- play()
attributes(n)

play <- function() {
symbols <- get_symbols()
structure(score(symbols), symbols = symbols)
}

three_play <- play()
three_play

slot_display <- function(prize){
# extract symbols
symbols <- attr(prize, "symbols")
# collapse symbols into single string
symbols <- paste(symbols, collapse = " ")
# combine symbol with prize as a regular expression
# \n is regular expression for new line (i.e. return or enter)
string <- paste(symbols, prize, sep = "\n$")
# display regular expression in console without quotes
cat(string)
}

slot_display(three_play)

class(three_play)
three_play

methods(print)
class(one_play) <- "slots"
class(
      one_play
)
attributes(one_play)
n <- 10000
class(n)
print.slots <- function(x, ...) {
cat("I'm using the print.slots method")
}
one_play
print.slots <- function(x, ...) {
cat("I'm using the print.slots method")
}

now <- Sys.time()
attributes(now)
attributes(now)
print.slots <- slot_display
one_play
now <- Sys.time()
attributes(now)

play <- function() {
symbols <- get_symbols()
structure(score(symbols), symbols = symbols, class = "slots")
}

attributes(play)
attributes(play())

methods(class = "numeric")
factor(play())
play1 <- play()
play2 <- play
playV <- c(play1, play2)
attributes(playV)
class(playV)